Synthesizing Quotient Lenses

### 2017-10-20 papers/QuotientLensSynth/main.tex
added related work section

\begin{figure}
  \centering
  \begin{subfigure}[b]{.49\textwidth}
    \centering
    \includegraphics{generated-graphs/times_opt}
    \caption{}
    \label{subfig:lenssize}
  \end{subfigure}
  \begin{subfigure}[b]{.49\textwidth}
    \includegraphics{times_new.eps}
    \caption{}
    \label{subfig:examplesused}
  \end{subfigure}
  \caption{Runtimes of Optician and \Name{}.
    In (a), we run Optician and \Name{} on Optician's benchmarks.  We find that
    there is only a negligible difference incurred by Optometrist's overhead.
    In (b), we run Optometrist on its benchmark suite.  We find that it is able
    to synthesize all quotient lenses in under 20 seconds, and typically
    finished in under 5 seconds. }
  \label{fig:times}
\end{figure}

@@

This paper also builds on our previous work where we showed how to synthesize
bijective lenses from a pair of regular expressions describing the source and
target types and a set of input-output examples. We refer the reader to the
Related Work section of \cite{popl18} for a survey of the area of
type-and-example driven synthesis.

>>>>

\begin{figure}
  \centering
  \begin{subfigure}[b]{.49\textwidth}
    \centering
    %\includegraphics{generated-graphs/times_opt}
    \caption{}
    \label{subfig:lenssize}
  \end{subfigure}
  \begin{subfigure}[b]{.49\textwidth}
    %\includegraphics{times_new.eps}
    \caption{}
    \label{subfig:examplesused}
  \end{subfigure}
  \caption{Runtimes of Optician and \Name{}.
    In (a), we run Optician and \Name{} on Optician's benchmarks.  We find that
    there is only a negligible difference incurred by Optometrist's overhead.
    In (b), we run Optometrist on its benchmark suite.  We find that it is able
    to synthesize all quotient lenses in under 20 seconds, and typically
    finished in under 5 seconds. }
  \label{fig:times}
\end{figure}

@@

As far as synthesis goes, though there is a good deal of recent research on
synthesizing unidirectional string
transformations~\cite{singh2012learning,le-pldi-2014,gulwani-popl-2014,
perelman2014test,Singh:blinkfill}, our system Optician, which we introduced in
~\cite{popl18}, is the first attempt to synthesize bidirectional
transformations that we are aware of. We compared Optician to two of these
unidirectional string transformers, Flash Fill~\cite{gulwani-popl-2014} and
FlashExtract~\cite{le-pldi-2014}, and found that these tools were unsuccessful
in synthesizing the complex transformations that can be expressed in Optician.
Furthermore, neither of these tools were able to infer transformations 
which occurred under two iterations.

Much of the research in synthesis assumes that the synthesizer is provided with
a collection of examples. Optician differs in that it requires that the
programmer supplies both examples {\em and} format descriptions in the form of
regular expressions. There are many other recent results showing how to
synthesize functions from type-based
specifications~\cite{augustsson-2004,osera+:pldi15,
feser-pldi-2015,scherer-icfp-2015,frankle+:popl16,armando+:pldi16}.
These systems enumerate programs of their target language, orienting their
search procedures to process only terms that are well-typed.
Optician is distinctive in that it synthesizes terms in a language with many
type equivalences. 
Perhaps the system most similar to Optician is InSynth~\cite{gvero-pldi-2013}, a
system for synthesizing terms in the simply-typed lambda calculus that addresses
equivalences on types.

Morpheus~\cite{morpheus} is another synthesis system that uses two
communicating synthesizers to generate programs.  In both Morpheus and
Optician, one synthesizer provides an 
outline for the program, and the other fills in that outline with program
details that satisfy the user's specifications.
This approach works well in large search spaces, which
require some enumerative search.
One important way in which Optician differs from Morpheus is that in
Morpheus, an outline is a sketch---an
\emph{expression}
containing holes---whereas
an outline in Optician is a pair of regular
expressions, i.e., a 
\emph{type}.  Moreover, in order to implement an efficient
search procedure, we had to create both a new type language and a new
term language for lenses.  Once we did so, we proved our new, more
constrained language
designed for synthesis was just as expressive as the original, more
flexible and compositional language designed for human programmers.
